Thesis: draft

### 2021-05-16 draft.tex
wrote conclusion

\chapter{Conclusions and outlook}
% also discuss need to investigate effect of changing N, S and p_act
\newpage

>>>>

\chapter{Conclusion}

Simple toy models were implemented to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions. Our goal was to get a better understanding of the impact of social media and algorithmic personalization on phenomena such as polarization and formation of echo chambers (opinion clusters) in society. \\
\newline
We found that topological correlations such as high clustering strongly enhance the formation of opinion clusters. Beside clustering, community structure also favors the formation of echo chambers. Real-world networks have a lower formation of echo chambers than network models with similar clustering and modularity. The heterogeneous contact pattern of these real-world networks is mostly likely the cause of this hampering. If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system. This is an intuitive result: if you shield people from opposing views and strengthen them in their own beliefs, it feels natural that the system will evolve towards a more polarized state. It is then expected that a more heterogeneous exposure to information reduces this polarization. We found that this is indeed the case if people are not stubborn. If, on the contrary, people have some level of stubbornness, we found the surprising result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure. This is an important result, since it may change our beliefs/assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media. More research on this topic is definitely needed in the near future. Finally, we found that distributing the opinions in communities instead of randomly distributing them may be hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
\newline
Our results strongly suggest that the effects of algorithmic personalization do not stand on their own. Instead, the interplay with certain network topologies may enhance or hamper these effects. Beside that, individual resistance strongly affects the impact of these filtering algorithms. It would therefore be incorrect to only point to algorithmic personalization as the cause of polarization. It is the interplay between many different aspects of the system (such as topological correlations, distribution of the opinions, individual resistance, content curration) that may potentially enhance or hamper the formation of opinion clusters.
% also discuss need to investigate effect of changing N, S and p_act
\newpage
